### 2014-11-12 evaluation.tex
Incorporated new edits

\begin{itemize}
\setlength{\itemsep}{2pt}
\item \textbf{LWW register}: A last-write-wins register that provides read
  and write operations. Each write is associated with a timestamp, which is
  used to resolve conflicting concurrent writes -- a newer write always
  wins.

@@

\item \textbf{Shopping list}: A collaborative shopping list that allows
  adding and deleting items.

@@

We compare the performance of different transaction isolation level choices
in Figure~\ref{grf:LWW-txn} using the LWW register. The numbers were
obtained under a 1DC configuration. The YCSB workload was modified to issue
10 operations per transaction, with a default 50:50 read:write mix. Each
operation is assumed to be eventual consistent. \cf{NoTxn} corresponds to a
configuration that does not use transactions. Compared to this, RC is only
12\% shower in terms of latency with 512 clients, whereas RR is 2.3X
slower. The difference between RC and \cf{NoTxn} is due to the meta-data
overhead of recording transaction information in the object state. For RR
transactions, the cost of capturing and maintaining a snapshot is the
biggest source of overhead.

>>>>

\begin{itemize}
\setlength{\itemsep}{2pt}
\item \textbf{LWW register}: A last-write-wins register that provides read
	and write operations, where the read returns the value of the latest write.

@@

\item \textbf{Shopping list}: A collaborative shopping list that allows
  concurrent addition and deletion of items.

@@

We compare the performance of different transaction isolation level choices in
Figure~\ref{grf:LWW-txn} using the LWW register. The numbers were obtained
under a 1DC configuration. The YCSB workload was modified to issue 10
operations per transaction, with a default 50:50 read:write mix. Each operation
is assumed to be eventually consistent. \cf{NoTxn} corresponds to a
configuration that does not use transactions. Compared to this, RC is only 12\%
shower in terms of latency with 512 clients, whereas RR is 2.3X slower. The
difference between RC and \cf{NoTxn} is due to the meta-data overhead of
recording transaction information in the object state. For RR transactions, the
cost of capturing and maintaining a snapshot is the biggest source of overhead.

### 2014-11-12 motivation.tex
Incorporated new edits

\noindent It takes a list of effects (the \emph{context} for the operation),
and an input argument, and returns a result along with an optional effect
(read-only operations return \cf{Nothing}). The new effect (if emitted) is
added to the state of the object at the current replica, and asynchronously
sent to other replicas. The implementation of the bank account operations in
\name is given in Figure~\ref{fig:ex}.

\begin{figure}
\begin{codehaskell}
data Acc = Deposit Int | Withdraw Int
				 | GetBalance

getBalance :: Operation Acc () Int
getBalance ctxt _ =
  let res = sum [x | Deposit x <- ctxt]
						- sum [x | Withdraw x <- ctxt]
	in (res, Nothing)

deposit :: Operation Acc Int ()
deposit _ amt = (amt, Just dollar Deposit amt)

withdraw :: Operation Acc Int Bool
withdraw ctxt v =
	if sel1 dollar getBalance ctxt () >= v
  then (True, Just dollar Withdraw v)
	else (False, Nothing)
\end{codehaskell}
\caption{Definition of a bank account expressed in Quelea.}
\label{fig:ex}
\end{figure}

The datatype \cf{Acc} represents the effect type for the bank account. The
context of the operations is a snapshot of the state of the object at some
replica. In this sense, every operation on the RDT is atomic, and thus
permitting sequential reasoning for implementing eventually consistent data
types. We have implemented a large corpus of RDTs for realistic benchmarks
including shopping carts, auction and micro-blogging sites in few tens of lines
of code.

@@

This notion of observable equivalence can be generalized to other RDTs as well.
For example, a last-writer-wins register with multiple updates is equivalent to
a register with only the last write. Similarly, a set with a collection of add
and remove operations is equivalent to a set with a series of additions of live
elements from the original set. Since the notion of observable equivalence is
specific to each RDT, we expect the programmer to provide a summarization
function (\cf{summarize}) of type \cf{[e] -> [e]} for each RDT. We utilize this
function to ensure that the state of an object remains summarized in each
replica. The summarization function for the bank account is {\footnotesize
\cf{summarize ctxt = [Deposit \$ sel1 \$ getBalance ctxt ()]}}.

@@

\begin{figure}[ht]
\centering
\subfigure[Unsafe withdraw]{\label{fig:unsafeWithdrawAnomaly}\includegraphics[width=0.34\columnwidth]{Figures/Motivation4}}
\hfill
\subfigure[Negative balance]{\label{fig:negativeBalanceAnomaly}\includegraphics[width=0.31\columnwidth]{Figures/Motivation2}}
\hfill
\subfigure[Missing update]{\label{fig:missingUpdateAnomaly}\includegraphics[width=0.26\columnwidth]{Figures/Motivation1}}
\caption{Anomalies possible under eventual consistency for the get balance operation.}
\label{fig:ba_anomalies}
\end{figure}

>>>>

\noindent It takes a list of effects (the \emph{history} of updates to that
object), and an input argument, and returns a result along with an optional
effect (read-only operations return \cf{Nothing}). The new effect (if emitted)
is added to the state of the object at the current replica, and asynchronously
sent to other replicas. The implementation of the bank account operations in
\name is given in Figure~\ref{fig:ex}.

\begin{figure}
\vspace{-2.3mm}
\begin{codehaskell}
data Acc = Deposit Int | Withdraw Int | GetBal

getBalance :: [Acc] -> () -> (Int, Maybe Acc)
getBalance hist _ =
  let res = sum [x | Deposit x <- hist]
						- sum [x | Withdraw x <- hist]
	in (res, Nothing)

deposit :: [Acc] -> Int -> ((), Maybe Acc)
deposit hist amt = ((), Just dollar Deposit amt)

withdraw :: [Acc] -> Int -> (Bool, Maybe Acc)
withdraw hist v =
	if sel1 dollar getBalance hist () >= v
  then (True, Just dollar Withdraw v)
	else (False, Nothing)
\end{codehaskell}
\caption{Definition of a bank account expressed in Quelea.}
\label{fig:ex}
\end{figure}

The datatype \cf{Acc} represents the effect type for the bank account. The
function \cf{sum} returns the sum of elements in the list, and \cf{sel1}
returns the first element of a tuple. For each operation, \cf{hist} is a
\emph{snapshot} of the state of the object at some replica. In this sense,
every operation on the RDT is atomic, and thus permitting sequential reasoning
for implementing eventually consistent data types. Here, \cf{getBalance} is a
read-only operation, \cf{deposit} always emits an effect, and \cf{withdraw}
only emits an effect if there is sufficient balance in the account. We have
implemented a large corpus of RDTs for realistic benchmarks including shopping
carts, auction and micro-blogging sites in few tens of lines of code.

@@

This notion of observable equivalence can be generalized to other RDTs as well.
For example, a last-writer-wins register with multiple updates is equivalent to
a register with only the last write. Similarly, a set with a collection of add
and remove operations is equivalent to a set with a series of additions of live
elements from the original set. Since the notion of observable equivalence is
specific to each RDT, we expect the programmer to provide a summarization
function (\cf{summarize}) of type \cf{[e] -> [e]} for each RDT, as a part of
the RDT specification.

The summarization function for the bank account is \cf{summarize hist =
[Deposit \$ sel1 \$ getBalance hist ()]}. Given a bank account history
\cf{hist}, the \cf{summarize} function returns a new history with a single
deposit of the current account balance. We utilize the summarization function
to ensure that the state of an object remains summarized in each replica.

@@

\begin{figure}[ht]
\vspace{-2mm}
\centering
\subfigure[Unsafe withdraw]{\label{fig:unsafeWithdrawAnomaly}\includegraphics[width=0.34\columnwidth]{Figures/Motivation4}}
\hfill
\subfigure[Negative balance]{\label{fig:negativeBalanceAnomaly}\includegraphics[width=0.31\columnwidth]{Figures/Motivation2}}
\hfill
\subfigure[Missing update]{\label{fig:missingUpdateAnomaly}\includegraphics[width=0.26\columnwidth]{Figures/Motivation1}}
\caption{Anomalies possible under eventual consistency for the get balance operation.}
\label{fig:ba_anomalies}
\end{figure}
